---
output_dir: "."
title: "RNotebookExample"
output: html_document
---

# This is a heading not a comment

```{r load-data, message=FALSE}
# load tidyverse library
library(tidyverse)
library(kableExtra) # used for kbl() 
library(rmarkdown) # used for paged_table()
library(ggpubr) # used for ggarrange()

# read file into R
pop_data <- read_csv("data/gapminder_data.csv")

# create a table with data from European countries in 2007 - orderd by life expectancy
euro_data_tbl <- pop_data %>%
                    filter(continent == "Europe" & year == 2007) %>%
                    select(-continent, -year) %>%
                    arrange(desc(lifeExp)) %>%
                    rename(Country = country, "Population Size" = pop, "Life Expectancy" = lifeExp, "GDP" = gdpPercap)

```

The results in euro_data_tbl are displayed in the Table below:

```{r kbl-table}
euro_data_tbl %>%
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>% 
          scroll_box(width = "100%", height = "200px")
```

This is a better way to display long tables:

```{r paged-table}
paged_table(euro_data_tbl)
```

# Adding Images

Next, the life expectancy in Europe is observed across different years.  The aim of this is to check if there was a change in life expectancy over the years.  The data used for the life expectancy over the years analysis is the gapminder data and the following data manipulation was performed:

```{r fig-data}
#keep on European data and change year to factor
euro_data_fig <- pop_data %>% 
                    filter(continent == "Europe")  %>% 
                    mutate(year=as_factor(year))

# keep only UK data
uk_data_fig <- pop_data %>%
                  filter(country == "United Kingdom") %>% 
                  mutate(year=as_factor(year))
```


The life expectancy for Europe over the years is plotted as following:

```{r fig, warning=FALSE, fig.cap="Life Expectancy in Europe over the years"}
#Euro plot
euro_plot <- euro_data_fig %>%
                ggplot(mapping=aes(x=year, y=lifeExp)) +
                geom_violin() +
                stat_summary(fun.y = median, geom = "point")

#draw euro plot
euro_plot

```

Now we will display more than one image together:

```{r fig-comp}

#UK plot
uk_plot <- uk_data_fig %>% 
              ggplot(mapping=aes(x=year, y=lifeExp)) +
              geom_point()

#plot EU and UK plots together
ggarrange(euro_plot, uk_plot, ncol = 2, nrow = 1, labels = "AUTO")

```





